---
title: "R's Internal Data Formats"
subtitle: ".Rda, .RData, .rds"
author: "StatistikinDD"
date: "Presentation Created: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["libs/_css/xaringan-themer.css", "libs/_css/my_css.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      ratio: 16:9
---

```{r setup, include = FALSE}

options(htmltools.dir.version = FALSE)
# knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, comment = "")
knitr::opts_chunk$set(echo = TRUE, comment = "")

```

# R's Internal Data Formats

## <br>

## .Rda and .RData

## How .rds is Different

## Comparing File Sizes to .csv

---

# The Diamonds Data (ggplot2)

.pull-left[
```{r plotly, eval = FALSE}

library(tidyverse)
library(plotly)

p <- ggplot(data = diamonds,
      aes(x = cut, fill = clarity)) +
      geom_bar(position = "dodge") +
      theme_bw()
p <- ggplotly(p)

library(widgetframe) #<<
frameWidget(p) #<<

```
]

.pull-right[
```{r, ref.label = "plotly", echo = FALSE, message = FALSE}
```

]
---

# Storing Data in .Rda <br>and .RData Format

```{r}
save(diamonds, file = "diamonds.Rda")
save(diamonds, file = "diamonds.RData")

my_fileinfo <- function(file) {
  file.info(file)$size |> utils:::format.object_size("auto")
}

my_fileinfo("diamonds.Rda")
my_fileinfo("diamonds.RData")

```

---

# Storing Several Objects in<br> .Rda / .RData Files

```{r}

save(diamonds, p, my_fileinfo, file = "diamonds.Rda")
rm(list = ls())
load("diamonds.Rda")
ls()

```

---

# .RData: Storing the <br>'Current Workspace'

This is what also happens interactively when you close down RStudio and confirm to save the Workspace Image.

```{r}

save.image()
rm(list = ls())
load(".RData")
ls()

```
## NB: There are good reasons NOT to rely on the Workspace Image!

---

# The .rds Format

## .rds Can Only Store ONE Object

```{r}
save(diamonds, file = "diamonds.Rda")
saveRDS(diamonds, file = "diamonds.rds")

my_fileinfo("diamonds.Rda")
my_fileinfo("diamonds.rds")

```

---

# Why I Prefer .rds over .Rda / .RData

```{r}

diamonds <- tibble(
  diamonds = c("Yellow", "Pink Champagne", "Synthetic"),
  assumption = c("Neutral", "Wife might like", "Doubtful")
)

diamonds_ggplot2 <- readRDS("diamonds.rds") #<<

load("diamonds.Rda") #<<

str(diamonds)

```

---

# Comparing File Sizes to .csv

```{r}

data.table::fwrite(diamonds, "diamonds.csv")

my_fileinfo("diamonds.csv")
my_fileinfo("diamonds.Rda")
my_fileinfo("diamonds.rds")

```

```{r, echo = FALSE}

unlink("diamonds.csv")
unlink("diamonds.Rda")
unlink("diamonds.RData")
unlink("diamonds.rds")

```

---

class: center, middle

# Thanks!

### Youtube: StatistikinDD

### Twitter: @StatistikinDD

### github: fjodor

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](https://yihui.org/knitr), and [R Markdown](https://rmarkdown.rstudio.com).

Thanks to **Yihui Xie** for *{knitr}* and *{xaringan}* and **Garrick Aden-Buie** for *{xaringanthemer}*.